Computing likelihoods under the SMC for a general class of ARGs.

### 2024-03-14 paper.tex
Remove whitespace

\begin{itemize}
    \item Models are good/essential: CwR useful but terrible for inference -> SMC
    \item Problem: we can't use models on inferred ARGs (Relate, tsinfer, ARGneedle). These methods produce point estimates. What we can't do:
        \begin{itemize}
            \item quantify uncertainty
            \item explore ARG space systematically around the point estimates. We don't have a metric on what an improvement might be.
            \item compare models of demography
            \item can't evaluate goodness of fit of model estimates
        \end{itemize}
    \item what we do here
\end{itemize}

>>>>

Model-based inference is a central element of population genetics.
The coalescent~\citep{kingman_genealogy_1982,kingman_coalescent_1982,hudson_testing_1983,
tajima_evolutionary_1983} has proved to be a useful model, and is the
basis of a wide variety of inferences [some citations?]. The incorporation
of recombination in to the coalescent process~\citep{hudson_properties_1983}
has proved to be much more difficult.
% FIXME rephrase, next two sentences, lifted from ARG paper
Early work on inference under the coalescent with recombination
focused on the problem of
inferring the parameters of the
stochastic process, where the ancestry was regarded as a
latent parameter to be averaged out
\citep[e.g.][]{griffiths_ancestral_1996,kuhner_maximum_2000,
nielsen_estimation_2000,
fearnhead_estimating_2001}.
% TODO reuse some of the ideas from the SMC paper here, it has
% some good explanations.
These methods met with limited success
because the state space of ARGs is overwhelmingly large, and
lacks a simple geometry or neighbourhood structure for inference or
sampling methods to  exploit.
The introduction of the Sequentially Markovian Coalescent (SMC)
approximation to the coalescent with recombination
by~\cite{mcvean_approximating_2005} was a seminal contribution,
and has made many different inferences
% TODO add some more recent stuff
feasible~\citep[e.g.]{li_inference_2011,paul_accurate_2011,schiffels_inferring_2014}.
In particular, the SMC is the basis for ARG sampling methods
like ARGweaver~\citep{rasmussen_genome-wide_2014}, which remains the
% TODO list other places ARGweaver has come out on top in terms of accuracy.
gold-standard in terms of accuracy~\citep{brandt2022evaluation}

\begin{itemize}
    \item Problem: we can't use models on inferred ARGs (Relate, tsinfer, ARGneedle). These methods produce point estimates. What we can't do:
        \begin{itemize}
            \item quantify uncertainty
            \item explore ARG space systematically around the point estimates. We don't have a metric on what an improvement might be.
            \item compare models of demography
            \item can't evaluate goodness of fit of model estimates
        \end{itemize}
    \item what we do here
\end{itemize}
